---
title: "Final Report"
subtitle: "Maternal Mortality, a fight for every women"
author: "Margaux Becue"
date: "2024-04-23"
format: html
theme: solar
execute:
  echo: false
  warning: false
  message: false
---
```{r}
#| label: gapminder-pop
#| include: false

#libraries
library(tidyverse)
library(plotly)
library(ggplot2)
library(viridisLite)


#data

unicef_indicator_2 <- read_delim("unicef_indicator_2.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
mapdata <- map_data("world") ##ggplot2

localisation <- read_delim("localisation.csv", 
                           delim = ";", escape_double = FALSE, trim_ws = TRUE)

md_region_2020 <- read_delim("md_region_2020.csv", 
                             delim = ";", escape_double = FALSE, trim_ws = TRUE)

unicef_indicator_2 <- read_delim("unicef_indicator_2.csv", 
                                 delim = ";", escape_double = FALSE, trim_ws = TRUE)
unicef_indicator_1 <- read_delim("unicef_indicator_1.csv", 
                                 delim = ";", escape_double = FALSE, trim_ws = TRUE)
Nigeria <- read_delim("Nigeria.csv", delim = ";", 
                      escape_double = FALSE, trim_ws = TRUE)

#transformations

mapdata <- full_join(mapdata,unicef_indicator_2)
data_2020 <- md_region_2020%>% full_join(localisation, by = c("region"="country"))
Law_data <- full_join(unicef_indicator_1,unicef_indicator_2)
```
---
 **Global Challenge**: Maternal deaths have decreased by 34% since 2000, but it's slow (2.1% annually). We're not on track to Sustainable Development Goals which is 6.4% annually before 2030. Troubling trend emerges: **setbacks** in regions between 2016 and 2020 (Western Europe, North America, Latin America and the Caribbean, as seen in the following table). Sub-Saharan Africa accounts for a staggering 70% (shown on Map). Terminology: *"woman"/"women"* encompasses all people with reproductive capacity.

```{r}
mapdata <- map_data("world") ##ggplot2

mapdata <- full_join(mapdata,unicef_indicator_2)

mapdatal<-mapdata%>% filter(!is.na(mapdata$maternal_deaths_estimated))
map1<-ggplot(mapdatal,aes(x=long, y=lat, group=group))+ geom_polygon(aes(fill=maternal_deaths_estimated), color="black")

map2 <- map1 + scale_fill_gradient(name = "% maternal_deaths_estimated", low = "violet", high =  "red", na.value = "grey50")+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y=element_blank(),
        axis.title.x=element_blank(),
        rect = element_blank())
map2
```

**Leading causes of Maternal mortality**: severe bleeding, high blood pressure, infections, and unsafe abortions, pre-existing health issues. Sudden and preventable: most deaths are preventable with skilled healthcare attendance during childbirth. Healthcare matters but can be double-edged: ***A well-functioning health system with enough resources and respectful treatment can significantly improve outcomes. Conversely, a poorly functioning system can make things worse.***

```{r}
data_2020 <- md_region_2020%>% full_join(localisation, by = c("region"="country"))


colnames(data_2020)[1] <- "country"

data_2020$country <- as.character(data_2020$country)
data_2020$country[data_2020$country == "USA"] <- "United States of America"
data_2020$country[data_2020$country == "UK"] <- "United Kingdom"
data_2020$country[data_2020$country == "Republic of Congo"] <- "Republic of the Congo"
data_2020$country[data_2020$country == "Democratic Republic of the Congo"] <- "Congo"

ggplot(data_2020) +
  aes(Region, maternal_deaths_estimated, color=Region) +
  geom_line()+
  theme(plot.title = element_text(size = rel(1)),
        strip.text = element_text(face="bold", size=9,lineheight=5.0), #title by column
        axis.text.x = element_blank()) #remove x-axis legend here
```

The concept of "embodiment" emphasizes that social and environmental factors are integrated into a woman's physical and mental health during pregnancy (ex: *Poverty can worsen existing health issues*).

```{r}
Law_data <- full_join(unicef_indicator_1,unicef_indicator_2)

plot(Law_data$maternal_deaths_estimated,Law_data$legal_frameworks_pourcentage,
     main= "Estimated number of deaths on pourcentage of legal frameworkds present", 
     xlab= "% Maternal deaths", ylab = "% of legal frameworks present")

abline(lm(Law_data$legal_frameworks_pourcentage~Law_data$maternal_deaths_estimated),
       col= "violet")

```

## Case Study of Nigeria

This case study focuses on Nigeria's *high maternal mortality rate*, which stood at **81747** in 2020, categorizing it as a stage 1 country in the Transition Model. Maternal death stems also from social and economic factors, as illustrated by the correlation between gender equity law enforcement and maternal deaths. ***We note Nigeria's low ranking in gender equality laws and highly high maternal death compared to countries like Australia, which have lower maternal mortality rates.***

```{r}
barplot(Nigeria$pourcentage_women_gave_birth_before_18,Nigeria$year,
        names.arg=c("2003","2008","2011","2013","2017","2018"), 
        ylab = "% women who gave birth before 18 years old", 
        col=Nigeria$pourcentage_women_gave_birth_before_18)

barplot(Nigeria$Percentage_deliveries_skilled_health_personnel,Nigeria$year,
        names.arg=c("2003","2008","2011","2013","2017","2018"), 
        ylab = "% deliveries by skilled health personel", 
        col=Nigeria$Percentage_deliveries_skilled_health_personnel)
```
## Hope persists 

Some countries have significantly reduced maternal mortality (annual MMR reductions exceeding 15%). In the case of Nigeria, we can note a **satisfying rise in deliveries attended by medical personnel and care for the babies post-natal and a decrease in the number of women having a child before 18 years**. The solution: every woman needs essential prenatal care, skilled birth attendance, and postnatal care. Investments, improved healthcare access, and strong partnerships are critical, especially in high-risk areas. Equity is the goal: ***We can create a world where all women have a safe pregnancy and childbirth by tackling healthcare disparities***.